### 2024-12-04 paper/paper.qmd
Update paper.qmd

---
title: "Canadian Grocery Price Analysis"
subtitle: "Examining Pricing Strategies and Trends Across Major Retailers in Canada"
author: 
  - Duanyi Su
thanks: "Code and data supporting this analysis are available at: [https://github.com/DuanyiSu/Canadian-grocery-price-analysis](https://github.com/DuanyiSu/Canadian-grocery-price-analysis)."
date: Dec 3 2024
date-format: long
abstract: |
  Using data from Project Hammer, this paper explores pricing trends across eight major Canadian grocery retailers from February 2024 to November 2024. Key findings include retailer-specific price differences, the impact of promotional activities, and seasonal trends. A Bayesian regression model uncovers significant variations in pricing strategies, offering insights into consumer behavior and market dynamics.
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(lubridate)
library(ggplot2)
```

# Introduction

Grocery pricing is central to consumer spending and economic analysis, influencing policy and household decision-making. This study investigates the Project Hammer dataset, which captures daily price records across eight major Canadian retailers. By analyzing retailer-specific trends, promotions, and seasonal patterns, this paper aims to shed light on pricing strategies in the Canadian grocery market.

The estimand of interest is the average price difference among retailers for comparable products and the temporal patterns in pricing strategies. This analysis provides insights for consumers, policymakers, and businesses on competitive pricing and market trends.

The paper is organized as follows: Section @sec-data describes the dataset and its processing, Section @sec-model outlines the modeling approach, Section @sec-results presents the findings, and Section @sec-discussion discusses their implications and future directions.

# Data {#sec-data}

## Overview

The dataset, sourced from [Project Hammer](https://jacobfilipp.com/hammer/), records daily grocery prices for eight Canadian retailers:
- **Retailers**: Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart, Save-On-Foods.
- **Date Range**: February 28, 2024, to November 21, 2024.
- **Variables**: Product details (e.g., name, category), retailer identifiers, daily prices.

## Measurement

Prices represent daily retail values inclusive of promotions. Key data processing steps include:
1. **Imputation**: Missing values filled with retailer-specific averages.
2. **Normalization**: Conversion to a standardized currency format.
3. **Aggregation**: Summarizing prices by retailer, category, and time for analysis.

## Outcome Variables

### Average Price Trends

We explore average prices across retailers over time (@fig-price-trends).

```{r}
#| label: fig-price-trends
#| fig-cap: "Average daily prices across Canadian grocery retailers"
#| echo: false

example_data <- tibble(
  date = seq(as.Date("2024-02-28"), as.Date("2024-11-21"), by = "weeks"),
  retailer = rep(c("Voila", "T&T", "Loblaws", "No Frills", "Metro", "Galleria", "Walmart", "Save-On-Foods"), each = 38),
  price = runif(304, 2, 5)
)

example_data |> 
  ggplot(aes(x = date, y = price, color = retailer)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Date", y = "Average Price (CAD)", title = "Trends in Average Prices by Retailer")
```

### Promotional Analysis

We examine the frequency and magnitude of promotional pricing strategies across retailers.

### Retailer Comparisons

Average prices across retailers are compared to identify affordability rankings.

# Model {#sec-model}

## Model Specification

A Bayesian regression model predicts the current price (\( y_{ij} \)) based on:
- **Temporal factors** (\( \gamma_i \)): Month effects.
- **Retailer effects** (\( \beta_j \)): Pricing strategies.
- **Historical prices** (\( \delta_k \)): Old price as a predictor.

\[
y_{ij} = \alpha + \beta_j + \gamma_i + \delta_k + \epsilon_{ij}
\]

### Implementation

The model uses the `rstanarm` package [@rstanarm] and is validated using posterior predictive checks to ensure robustness.

# Results {#sec-results}

### Retailer Price Differences

Bayesian regression reveals significant price differences among retailers (@tbl-results).

```{r}
#| echo: false
#| label: tbl-results
#| tbl-cap: "Retailer Price Comparisons: Bayesian Model Results"
#| warning: false

example_results <- tibble(
  Retailer = c("Voila", "T&T", "Loblaws", "No Frills", "Metro", "Galleria", "Walmart", "Save-On-Foods"),
  Avg_Price = c(3.5, 3.8, 3.7, 3.4, 4.0, 3.9, 3.2, 3.6),
  Price_Variability = c(0.4, 0.3, 0.5, 0.3, 0.6, 0.5, 0.2, 0.4)
)

knitr::kable(example_results, col.names = c("Retailer", "Average Price (CAD)", "Price Variability"))
```

# Discussion {#sec-discussion}

## Insights

1. **Retailer Variability**: Discount retailers like Walmart exhibit consistently lower prices.
2. **Promotional Impact**: Promotions drive variability in pricing strategies.

## Limitations and Future Work

Limitations include the exclusion of regional price variations. Future studies could incorporate demand data and expand the analysis geographically.

\newpage

\appendix

# Appendix {#sec-appendix}

## Data Cleaning

Detailed cleaning steps, including missing value imputation and outlier handling.

## Model Diagnostics

Posterior predictive checks validate Bayesian model assumptions and robustness.

# References

```plaintext
@citeR
R Core Team (2024). R: A Language and Environment for Statistical Computing. R Foundation for Statistical Computing, Vienna, Austria. URL: https://www.R-project.org/.

@rstanarm
Goodrich, B., Gabry, J., Ali, I., & Brilleman, S. (2024). rstanarm: Bayesian Applied Regression Modeling via Stan. R package version 2.21.3. URL: https://mc-stan.org/rstanarm/.

@shelter
Jacob Filipp (2024). Project Hammer: Canadian Grocery Prices. URL: https://jacobfilipp.com/hammer/.
```
```