### 2024-12-04 other/datasheet/datasheet.qmd
Update datasheet.qmd

---
title: "Datasheet for 'Canadian Grocery Prices'"
subtitle: "The Project Hammer Dataset"
author: 
  - Duanyi Su
thanks: "Code and data supporting this analysis are available at: [https://github.com/DuanyiSu/Canadian-grocery-price-analysis](https://github.com/DuanyiSu/Canadian-grocery-price-analysis)."
date: "December 3, 2024"
date-format: long
abstract: |
  This paper explores pricing trends across eight major Canadian grocery retailers from February 2024 to November 2024 using data from Project Hammer. Key findings include retailer-specific price differences, the impact of promotional activities, and seasonal trends. Bayesian regression models reveal significant variations in pricing strategies, offering insights into consumer behavior and market dynamics.
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled?*  
   The dataset was created as part of Project Hammer to analyze competition and pricing in the Canadian grocery sector. It provides historical pricing data across leading retailers, filling the gap in publicly available structured information for academic and policy research.

2. *Who created the dataset?*  
   Jacob Filipp, as part of Project Hammer, created this dataset to support transparency and policy analysis in grocery pricing.

3. *Who funded the dataset creation?*  
   The dataset was self-initiated by Jacob Filipp without explicit external funding.

**Composition**

1. *What do the instances represent?*  
   Instances include product metadata (e.g., name, brand, vendor) and time-series price data (e.g., current price, historical price).

2. *How many instances are there?*  
   Thousands of instances spanning February to November 2024, covering eight major Canadian grocery retailers.

**Collection Process**

1. *How was the data collected?*  
   Data was scraped from publicly available grocery platforms using software tools and cleaned to ensure consistency.

2. *Were any ethical review processes conducted?*  
   None, as the dataset contains publicly accessible information without private or sensitive data.

**Uses**

1. *Has the dataset been used for other tasks?*  
   Yes, it has been used for analyzing retailer-specific pricing trends, seasonal variations, and promotional impacts in grocery pricing.

2. *What other tasks could the dataset be used for?*  
   Time-series forecasting, consumer behavior studies, and competition analysis.

3. *Are there tasks for which the dataset should not be used?*  
   It should not be used to infer comprehensive market behavior as it excludes local and small-scale retailers.

**Preprocessing**

1. *What preprocessing was performed?*  
   Data was filtered to include Walmart and T&T beef-related products, removing non-relevant fields. Date information was cleaned, and missing values were handled.

2. *Is the raw data available?*  
   Yes, available at [Project Hammer](https://jacobfilipp.com/hammer/).

**Distribution**

1. *How will the dataset be distributed?*  
   The dataset is publicly available via the Project Hammer website and GitHub repository.

2. *Will the dataset be updated?*  
   Updates are expected periodically, managed by Jacob Filipp.

**Maintenance**

1. *Who maintains the dataset?*  
   Jacob Filipp supports and maintains the dataset through Project Hammer.

2. *How can users contact the maintainer?*  
   Via email at jacob@jacobfilipp.com or GitHub repository issues.

**References**
```{bibliography}
```